---
title: "IRS 527 Filings Analysis"
author: "Isaac Flath"
date: '2022-02-01T21:13:14-05:00'
output:
  html_document:
    number_sections: no
    fig_caption: yes
    toc: yes
    toc_float: yes
    fig_width: 7
    fig_height: 4.5
    theme: cosmo
    highlight: tango
    code_folding: hide
categories: Politics
tags:
- IRS
- 527 Orgs
---

```{css custom_css, echo = FALSE}
#header {
    background-image: linear-gradient(120deg, lightblue, blue);
}
.title.toc-ignore {
  font-weight: 1000;
}
h1{
    background-image: linear-gradient(120deg, lightblue, blue);
}
```

```{r knitr_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

knitr::opts_chunk$set(out.width = "100%", fig.height=6, split=FALSE, fig.align = 'default')
options(dplyr.summarise.inform = FALSE)
```

# Introduction

# Setup {.tabset .tabset-fade .tabset-pills}

## Imports

These are all standard imports that I use. I haven't filtered out ones not used in this particular report, but these are my standard imports I start with on projects.

```{r imports, results='hide'}
library(DBI)

# general visualisation
library('ggplot2') # visualisation
library('scales') # visualisation
library('patchwork') # visualisation
library('RColorBrewer') # visualisation
library('corrplot') # visualisation
library('ggthemes') # visualisation
library('viridis') # visualisation
library('gt') # table styling

# general data manipulation
library('dplyr') # data manipulation
library('readr') # input/output
library('vroom') # input/output
library('tibble') # data wrangling
library('tidyr') # data wrangling
library('stringr') # string manipulation
library('forcats') # factor manipulation
library('janitor') # cleaning
library('lubridate') # Date wrangling


# specific
library('glue') # encoding
library('naniar') # missing values
#library('gganimate') # visualisation
library('ggrepel') # visualisation
library('grid') # visualisation
library('GGally') # visualisation
library('ggforce') # visualisation
#library('ggtext') # visualisation
library(rlang) # eval string as symbols
library(lazyeval) # eval string as symbols
```

## Paths

```{r}

```

## UDFs

```{r udf}
`%notin%` <- Negate(`%in%`)
```

## Connect to DB

Here we connect to our sqlite database.  This database was created in [a previous article](https://rpubs.com/iflath/irs-527-filings), so you will need to use code from there if you want to recreate this post.

```{r}
db_path <- "~/code_outputs/irs-527-filings/irs.sqlite"

con <- dbConnect(RSQLite::SQLite(), db_path)
tables <- dbListTables(con)
```












